How does the latest Prophet forecasting model compare to the ARIMA model? Evaluation of BTC/ZAR predictability through a Mincer-Zarnowitz approach

### 2017-09-14 Draft/Write-up/Draft.tex
Final Draft. Sigh

---
title: "How does the latest Prophet forecasting model compare to the ARIMA model? Evaluation of BTC/ZAR predictability through a Mincer-Zarnowitz approach"
Author: "Fiona Ganie"  
Student: "GNXFIO001"
bibliography: Tex/ref.bib       
output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 5 
    fig_height: 4
abstract: |
  The Autoregressive Integrated Moving Average (ARIMA) model is one of the most widely used time series models that has gained popularity in the exchange rate market due to its ease of implementation and tractability. With the evolution of computational power, soft computing techniques have since been used in exchange rate markets. The ARIMA model has been used as the standard benchmark model against which these complex methods have been compared. Although they have produced significantly better results than the ARIMA model, these models lack interpretability and building them is a challenging task. Prophet is a sophisticated model that differs to traditional time series models in that it can be utilised by non-experts who have little knowledge about the statistical intricacies involved in the model, however have domain knowledge about the data generating process. Prophet is also robust to outliers and can handle missing values in the time series without the need for interpolation. This paper compares the out-of-sample forecasts of Prophet and ARIMA over varying forecast horizons, using a Mincer-Zarnowitz approach for forecast evaluation. A complete dataset of the prices of Bitcoin/Rand as well as a dataset which has missing prices and outliers present is used in the analysis. It is found that the ARIMA model produces the most efficient and unbiased forecasts over all forecast horizons, and that the ARIMA model is more robust when faced with missing values and outliers present in the data. 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')

```

# Introduction \label{Introduction}
The Autoregressive Integrated Moving Average (ARIMA) model is one of the most widely used time series models that has attracted attention in financial market forecasting [@khashei2009]. Although the Random Walk model has typically been applied to foreign exchange markets and has produced superior results, some researchers have contended that foreign exchange markets are not efficient and believe that future prices depend on current and past events [@abu1996]. This has led to the application of the ARIMA model to exchange rate problems, where it has since gained popularity due to its ease of implementation and tractability. The ARIMA model is easy to use and yields good forecasts over short forecast horizons, however the linearity of the ARIMA model fails to adequately capture the non-linearity inherent in exchange rate data [@zhang1998].

@@

In December 2016, Facebook open-sourced their forecasting model Prophet. Prophet differs to traditional time series model such as the ARIMA, in that it can produce high quality forecasts in a straightforward way. Prophet is a sophisticated model that provides informative results, however is configurable and easy to use.  It can be utilised by non-experts who have little knowledge about the statistical intricacies involved in the model, however have domain knowledge about the data generating process. This knowledge can be easily incorporated into the model through its intuitively adjustable parameters [@taylor2017]. Furthermore, Prophet is robust to outliers and can handle missing values in the time series without the need for interpolation [@taylor2017]. This allows analysts without knowledge on how to pre-process the data, to utilise the model without sacrificing predictive accuracy.  If Prophet produces forecasts that are as good as the ARIMA model, it can be compared to more complex forecasting methods that are less tractable and flexible.

@@

1. Compare the out-of-sample forecasts of ARIMA and Prophet over varying forecast horizons.

2. Compare the out-of-sample forecast performance with missing values and outliers present in the data.

Section \ref{Background and Theory} will present the key findings from the major works in which comparisons have been made to the ARIMA model in exchange rate forecasting, as well as the theory behind the ARIMA and Prophet model. Section \ref{Data} will provide a brief explanation and motivation for the dataset used in this paper. Section \ref{Methodology} will describe the methodology used in selecting the ARIMA and Prophet models and the Mincer-Zarnowitz approach of measuring forecast accuracy. The results from applying the ARIMA and Prophet model to the dataset is presented in Section \ref{Results}. Finally, Section \ref{Discussion and Conclusions discusses the results in a statistical context.

# Background and Theory \label{Background and Theory}

## Forecasting Exchange Rates with the ARIMA model
ARIMA models became highly popular since its introduction by Box and Jenkins, when it was shown that they could outperform complex econometric models in a variety of situations [@spyros1997]. The ARIMA model expresses the process {$y_t$} as a function of the weighted average of past values of the process and lagged values of the residuals. The weighted average of the past p values of the process represents an autoregressive (AR) process of order p. It feeds back past values of the process into the current value, inducing correlation between all lags of the process. The weighted average of the q lagged residuals represents a moving average (MA) process of order q. The purpose of mixing the MA process with the AR process is to reduce the large number of past values required by AR processes and to control for the autocorrelation which it creates between lagged values of the process. The combination of the AR(p) and MA(q) process results in a more parsimonious model, and forms a stationary autoregressive moving average (ARMA(p,q)) process defined as:

$$ 
	y_t = c + \sum_{i = 1}^{p} \theta_{i} y_{t-i} + \sum_{i = 1}^{q} \phi_{i} e_{t-i}+ e_t \label{eq1} \\ \notag
$$
where c is a constant and {$e_t$} is a white noise process with zero mean and variance $\sigma^2$.

The ARIMA(p,d,q) model generalises the ARMA model in that it includes both stationary and non-stationary processes. The parameter d is the degree of differencing required to render the process stationary. If d is equal to zero the process is stationary and equivalent to an ARMA model, and if d is strictly positive the process requires differencing to make it stationary. The ARIMA model can be defined succinctly using the backward shift operator B, which shifts the process back by one unit of time, and is defined as $By_t = y_{t-1}$. The ARIMA model has the form:

$$ 
	(1-\sum_{i = 1}^{p} \theta_{i} B^i)(1-B)^dy_t = c + (1 + \sum_{i = 1}^{q} \phi_{i} B^i)e_t \label{eq2} \\ \notag 
$$
where c is a constant and {$e_t$} is a white noise process with zero mean and variance $\sigma^2$.

ARIMA models have commonly been used in financial forecasting and are popular for observing stock prices and exchange rates due to its power and statistical properties [@lin2012]. They have frequently been used as a benchmark to compare new forecasting techniques that have emerged over time. @kamruzzaman2003 applied an ARIMA model to forecast the exchange rate between the Australian dollar and six other currencies, and used the forecast errors as well as the accuracy of the direction of the forecasts to evaluate the performance of three Neural Network models. In a similar study, @khashei2012 compared the predictive capability of their proposed hybrid model consisting of an ARIMA and a Probabilistic Neural Network (PNN) to the traditional ARIMA model, and justified the use of the ARIMA as a benchmark by claiming that it is the most important linear model. Their comparison was made by investigating the forecasts between the British pound and US dollar over varying forecast horizons. Not only has the ARIMA model been useful as a benchmark, but it has also yielded satisfactory results when predicting exchange rates. When @nwankwo2014 forecasted the rate between the Nairo and dollar, diagnostic testing revealed that the ARIMA(1,0,0) model was the best fit for the data based on Akaike’s Information Criterion (AIC).

The ARIMA model is attractive as it is tractable and produces good short-term forecasts when more than 100 observations are used [@tseng2001]. Although the ARIMA model has the advantage of ease of implementation and flexibility, it fails to capture the non-linearity and volatility present in exchange rate data. Over time, the ARIMA model has evolved to cater for a wider variety of data and to compensate for some of its shortcomings. The most popular versions of the ARIMA model that has been implemented in exchange rate forecasting is the Seasonal ARIMA (SARIMA) and Fractional ARIMA model (FARIMA).

The SARIMA model was introduced to capture the periodic behaviour of data and extends the ARIMA class by including a term for seasonal differencing. @etuk2013 modelled the Naira/CFA Franc exchange rate which exhibited monthly seasonality using an additive SARIMA model, to demonstrate that it can be a useful fit for exchange rate data which displays seasonality. Their results showed that the SARIMA model adequately described the variation in the exchange rate series. The ARFIMA model generalises the ARIMA model in that the degree required to make the data stationary can assume any real value, and is no longer restricted to the integer domain. The ARFIMA model has the ability to capture the dependence between observations that are widely spread apart in time [@cheung1993]. This makes the model more parsimonious since it can capture long memory in data as well as short term dynamics [@cheung1993]. @cheung1993 fitted the ARFIMA model to examine five exchange rates, and found that there was strong evidence of long memory in the exchange rate time series.

Generalised autoregressive conditional heteroskedasticity (GARCH) models were later developed in an attempt to capture the volatility in financial markets [@anastasakis2009]. Hsieh (1989) applied a GARCH model to investigate five exchange rates and his results showed that although the GARCH model outperformed the random walk, some non-linear information still remained in the residuals.

## Other techniques used to forecast exchange rates
Over time, financial forecasting methods have moved away from linear models like ARIMA and GARCH, to soft computing techniques. These more complex techniques are non-linear and can fit complex time series more easily [@castillo2002]. Unlike the ARIMA model, soft computing techniques do not impose structural assumptions on the model apriori [@castillo2002]. Some of the most commonly used artificial intelligence methods used to forecast exchange rate data are Neural Networks and Fuzzy Logistic Systems.

Artificial Neural Networks (ANNs) have had many successful applications in forecasting exchange rates and are more advantageous than other non-linear forecasting methods. They are data driven, can adapt to non-stationary environments and can approximate any continuous function [@khashei2011]. In a study done by @fahimifard2009, the ANN was found as an effective way to improve the forecasts of exchange rates. Superior results were produced when compared to the ARIMA and GARCH model using the root mean square error (RMSE), mean square error (MSE) and mean absolute difference as a measure of performance. Fuzzy Logistic Systems (FLSs) were initially developed to solve problems involving linguistic terms, and have been successfully used in financial forecasting [@khashei2009]. Fuzzy logic tries to imitate human reasoning and the decision-making process and allows for finer rather than discrete decisions to be provided. @santos2007 investigated how well FLSs and ANNs perform compared to the traditional ARMA and GARCH model. He examined the forecasts of Brazilian exchange rate returns by considering different frequencies of the series and comparing their one step-ahead forecasts. By analysing accuracy statistics, he found that FLSs and ANNs achieved higher returns based on the forecasts they produced. Similar results were found by @khashei2009 when he analysed the predictive capabilities of FLSs, ANNs, the traditional ARIMA model, and a Fuzzy ARIMA model.

Although ANNs have been broadly applied in financial forecasting, the process of building them is a complex task and there is no consistent method of design compared to the traditional Box-Jenkins ARIMA model. Unlike ARIMA models, the performance of ANNs is sensitive to many modelling factors such as the number of input nodes included and the size of the training sample chosen [@zhang1998]. Like ANNs there is no systematic approach to designing FLSs and they are only understandable when simple. Although FLSs has the advantage over ARIMA models that they can be applied to data with few observations available, it gives acceptable rather than accurate results and are more suitable for problems which do not require high accuracy.

## Hybrid ARIMA Models
Over time, many researchers began to think of ways in which to harness the advantages of tractable linear models such as the ARIMA model, and of more complex non-linear models. By combining the ARIMA model with other forecasting methods, the advantages of both forecasting methods are leveraged while simultaneously improving their limitations. Some of the hybrid models which have commonly been used in exchange rate forecasting are the Fuzzy ARIMA and ANN-ARIMA model.

The ARIMA model produces very accurate forecasts over short time horizons however it has the limitation of requiring more than 100 observations of historical data to yield accurate results [@tseng2001]. In a world that is constantly changing and with the rapid advancement of technology, access to large amounts of historical data is difficult to obtain. On the contrary, a fuzzy regression model requires little historical data however produces wide prediction intervals if extreme values are present in the data. The Fuzzy ARIMA model combines the ARIMA and fuzzy regression model to exploit the advantages of both models while simultaneously overcoming their limitations. @tseng2001 proposed applying a Fuzzy ARIMA model to forecast the exchange rate of Taiwan dollars to US dollars to demonstrate the model’s appropriateness and power. The Fuzzy ARIMA not only produced forecasts that were superior to the ARIMA and fuzzy time series models but also provided an upper and lower bound which can be used by decision makers to determine the best and worst possible situations.

Although ARIMA models are powerful, they require non-stationary data to be differenced and impose prior assumptions onto the distribution of the data [@ince2006]. In contrast, machine learning techniques such as ANNs do not impose any assumptions onto the data generating process however, being data-driven, are sensitive to the number of input nodes used. The ANN-ARIMA model draws on the strengths of both models and overcomes these individual difficulties. @ince2006 created an ANN-ARIMA model by using the ARIMA model to determine the number input nodes required by an ANN for three exchange rates. When the ANN-ARIMA model was compared to the pure ARIMA model, the hybrid model outperformed the ARIMA based on the MSE.

Although the traditional ARIMA models produce less superior forecasts than its hybrid forms and other complex non-linear techniques, its forecasts are still satisfactory. They are simple models that are easy to implement and have a consistent method of model design and selection. ARIMA models are also more robust and efficient than complex structural models in relation to short-run forecasting. The fact that they have been used as the foundation for more advanced models and have commonly been used as a benchmark for comparison, justifies it as a good starting point to compare it to Facebook’s forecasting method, Prophet, that was recently released.

## Forecasting with Prophet 
The techniques that have been considered for exchange rate forecasting thus far, require the analyst to have vocational knowledge about time series. Prophet differs to traditional time series models in that it is flexible and can be customised by a large number of non-experts who have little knowledge about time series, however have domain knowledge about the data generating process. Prophet allows for a large number of forecasts to be produced across a variety of problems and consists of a robust evaluation system that allows for a large number of forecasts be evaluated and compared. This is Facebook’s definition of forecasting at scale.
Prophet is similar to a Generalized Additive Model (GAM) - an additive regression model that consists of non-linear and linear regression functions applied to predictor variables [@taylor2017]. The decomposable model is of the form:

$$ 
	y(t) = g(t) + s(t) + h(t) + e_t
$$
where the components of the model represent the growth, seasonality and holiday respectively, and $e_t$ is white noise.

Prophet, like the GAM, frames the forecasting problem as a curve fitting exercise and uses backfitting to find the regression functions. This allows for the model to be fitted quickly and missing values and large outliers to be handled elegantly. The regression model also provides model flexibility and allows the analyst to interactively change model parameters [@taylor2017].

The growth component of Prophet may be modelled as a linear or non-linear function of time. Linear growth is modelled by a piecewise constant function while non-linear growth is modelled similar to population growths which use a logistic growth model [@taylor2017]. A time-varying upper limit may be specified for logistic growth, at which point the forecasts will saturate. This carrying capacity allows the analyst to incorporate their prior knowledge about the maximum obtainable growth level such as the total market or population size into the model. Prophet accounts for changes in the trajectory of this trend by automatically detecting and selecting changepoints in the data at which the growth rate is allowed to change. These changepoints have a Laplace prior distribution placed on them and its scale parameter may be used to adjust the flexibility of the trend and to choose how aggressively the model should follow historical trend changes. Analysts may also adjust the number of potential changepoints included or manually specify their location. This allows non-experts with knowledge about events that may affect the growth rate to use the parameter as a knob to either increase or decrease the number of changepoints included [@taylor2017]. It also allows for the analyst to add changepoints which the automatic selection procedure may have missed or remove changepoints when the model is overfitting historical trends [@taylor2017].

The decomposable form of the model allows for multiple seasonality components with different periods to be added to the model. Seasonality components are modelled by a Fourier series and have a Normal prior distribution placed on its parameters [@taylor2017]. The spread parameter can be adjusted by analysts to smooth the seasonality and change how much of historical seasonality is projected into the future [@taylor2017].

The analyst may also provide a list of important events and holidays which have impacted the time series in the past or which they know might impact it in the future. The list could include the name, date, and country in which they have taken place or are expected to take place [@taylor2017]. By specifying the country of occurrence, separate lists can be populated for global events and holidays, and country-specific events and holidays. The union of the global and country-specific lists can then be used for forecasting. Like seasonality, a Normal prior distribution is placed on the parameters of the holiday component, and the scale parameter can be adjusted by analysts to smooth the holidays [@taylor2017].

Prophet’s Bayesian approach to forecasting allows the analyst to incorporate their expert knowledge into the model building process and has produced significantly improved forecasts compared to the ARIMA model. @taylor2017 forecasted the number of events on Facebook using Prophet. The time series was impacted by holidays, had strong multi-period seasonality, and a piecewise trend. The forecasts produced by Prophet were compared to common forecasting techniques such as exponential smoothing, ARIMA, the seasonal naïve, and the naïve model. While exponential smoothing and the seasonal naïve model were quite robust, the ARIMA forecasts were fragile. No model besides Prophet accounted for the dips around holidays and the upward trend of the time series towards later observations.

Prophet’s ability to forecast at scale enables it to model a wide variety of data and may be able to adequately fit exchange rate data. Its non-linear components could capture the non-linearity present in exchange rates. In contrast, the ARIMA model is a linear function of previous observations and lagged residuals and fails to capture the non-linearity inherent in exchange rate data. Prophet performs well on data with strong multiple “human scale” seasonalities and historical trend changes. This differs to the ARIMA model which requires the data to be de-trended and the variance stabilised before the model can be fitted. Hence Prophet could model the weekly seasonality of closing prices due to low trading activity which occurs around the weekend and high trading activity which occurs mid-week.

>>>>

\documentclass[12pt,a4paper]{article}

\usepackage{url}
\usepackage{appendix}
\usepackage[british]{babel}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{graphicx}
\numberwithin{equation}{section}
\numberwithin{figure}{section}
\numberwithin{table}{section}

\usepackage[round]{natbib}
\bibliographystyle{natbib}
\def\bibsection{\section*{References}}

% Wrap around which gives all figures included the [H] command, or places it "here". This can be tedious to code in Rmarkdown.
\usepackage{float}
\let\origfigure\figure
\let\endorigfigure\endfigure
\renewenvironment{figure}[1][2] {
    \expandafter\origfigure\expandafter[H]
} {
    \endorigfigure
}

\let\origtable\table
\let\endorigtable\endtable
\renewenvironment{table}[1][2] {
    \expandafter\origtable\expandafter[H]
} {
    \endorigtable
}

\def\tightlist{}

\bibpunct[:]{(}{)}{;}{a}{,}{,}

\renewcommand{\baselinestretch}{1.5}

\begin{document}

\begin{titlepage}

\begin{center}
{\Huge \bf How does the latest Prophet forecasting model compare to the ARIMA
model? Evaluation of BTC/ZAR predictability through a Mincer-Zarnowitz
approach}\\
\today\\
Fiona Ganie (GNXFIO001)\\
{\tt GNXFIO001@myuct.ac.za}
\end{center}

\begin{abstract}
The Autoregressive Integrated Moving Average (ARIMA) model is one of the
most widely used time series models that has gained popularity in the
exchange rate market due to its ease of implementation and tractability.
With the evolution of computational power, soft computing techniques
have since been used in exchange rate markets. The ARIMA model has been
used as the standard benchmark model against which these complex methods
have been compared. Although they have produced significantly better
results than the ARIMA model, these models lack interpretability and
building them is a challenging task. Prophet is a sophisticated model
that differs to traditional time series models in that it can be
utilised by non-experts who have little knowledge about the statistical
intricacies involved in the model, however have domain knowledge about
the data generating process. Prophet is also robust to outliers and can
handle missing values in the time series without the need for
interpolation. This paper compares the out-of-sample forecasts of
Prophet and ARIMA over varying forecast horizons, using a
Mincer-Zarnowitz approach for forecast evaluation. A complete dataset of
the prices of Bitcoin/Rand as well as a dataset which has missing prices
and outliers present is used in the analysis. It is found that the ARIMA
model produces the most efficient and unbiased forecasts over all
forecast horizons, and that the ARIMA model is more robust when faced
with missing values and outliers present in the data.
\noindent
Keywords: Prophet, ARIMA, Mincer-Zarnowitz,   Box-Jenkins, Bitcoin.
\end{abstract}
\end{titlepage}

\pagenumbering{arabic}

\section{\texorpdfstring{Introduction
\label{Introduction}}{Introduction }}\label{introduction}

The Autoregressive Integrated Moving Average (ARIMA) model is one of the
most widely used time series models that has attracted attention in
financial market forecasting (Khashei, Bijari, and Ardali 2009).
Although the Random Walk model has typically been applied to foreign
exchange markets and has produced superior results, some researchers
have contended that foreign exchange markets are not efficient and
believe that future prices depend on current and past events
(Abu-Mostafa and Atiya 1996). This has led to the application of the
ARIMA model to exchange rate problems, where it has since gained
popularity due to its ease of implementation and tractability. The ARIMA
model is easy to use and yields good forecasts over short forecast
horizons, however the linearity of the ARIMA model fails to adequately
capture the non-linearity inherent in exchange rate data (Zhang and Hu
1998).

@@

In December 2016, Facebook open-sourced their forecasting model Prophet.
Prophet differs to traditional time series model such as the ARIMA, in
that it can produce high quality forecasts in a straightforward way.
Prophet is a sophisticated model that provides informative results,
however is configurable and easy to use. It can be utilised by
non-experts who have little knowledge about the statistical intricacies
involved in the model, however have domain knowledge about the data
generating process. This knowledge can be easily incorporated into the
model through its intuitively adjustable parameters (Taylor and Letham
2017). Furthermore, Prophet is robust to outliers and can handle missing
values in the time series without the need for interpolation (Taylor and
Letham 2017). This allows analysts without knowledge on how to
pre-process the data, to utilise the model without sacrificing
predictive accuracy. If Prophet produces forecasts that are as good as
the ARIMA model, it can be compared to more complex forecasting methods
that are less tractable and flexible.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Compare the out-of-sample forecasts of ARIMA and Prophet over varying
  forecast horizons.
\item
  Compare the out-of-sample forecast performance with missing values and
  outliers present in the data.
\end{enumerate}

Section \ref{Background and Theory} will present the key findings from
the major works in which comparisons have been made to the ARIMA model
in exchange rate forecasting, as well as the theory behind the ARIMA and
Prophet model. Section \ref{Data} will provide a brief explanation and
motivation for the dataset used in this paper. Section \ref{Methodology}
will describe the methodology used in selecting the ARIMA and Prophet
models and the Mincer-Zarnowitz approach of measuring forecast accuracy.
The results from applying the ARIMA and Prophet model to the dataset is
presented in Section \ref{Results}. Finally, Section
\textbackslash{}ref\{Discussion and Conclusions discusses the results in
a statistical context.

\section{\texorpdfstring{Background and Theory
\label{Background and Theory}}{Background and Theory }}\label{background-and-theory}

\subsection{Forecasting Exchange Rates with the ARIMA
model}\label{forecasting-exchange-rates-with-the-arima-model}

ARIMA models became highly popular since its introduction by Box and
Jenkins, when it was shown that they could outperform complex
econometric models in a variety of situations (Hibon and Makridakis
1997). The ARIMA model expresses the process \{\(y_t\)\} as a function
of the weighted average of past values of the process and lagged values
of the residuals. The weighted average of the past p values of the
process represents an autoregressive (AR) process of order p.~It feeds
back past values of the process into the current value, inducing
correlation between all lags of the process. The weighted average of the
q lagged residuals represents a moving average (MA) process of order q.
The purpose of mixing the MA process with the AR process is to reduce
the large number of past values required by AR processes and to control
for the autocorrelation which it creates between lagged values of the
process. The combination of the AR(p) and MA(q) process results in a
more parsimonious model, and forms a stationary autoregressive moving
average (ARMA(p,q)) process defined as:

\[ 
    y_t = c + \sum_{i = 1}^{p} \theta_{i} y_{t-i} + \sum_{i = 1}^{q} \phi_{i} e_{t-i}+ e_t \label{eq1} \\ \notag
\] where c is a constant and \{\(e_t\)\} is a white noise process with
zero mean and variance \(\sigma^2\).

The ARIMA(p,d,q) model generalises the ARMA model in that it includes
both stationary and non-stationary processes. The parameter d is the
degree of differencing required to render the process stationary. If d
is equal to zero the process is stationary and equivalent to an ARMA
model, and if d is strictly positive the process requires differencing
to make it stationary. The ARIMA model can be defined succinctly using
the backward shift operator B, which shifts the process back by one unit
of time, and is defined as \(By_t = y_{t-1}\). The ARIMA model has the
form:

\[ 
    (1-\sum_{i = 1}^{p} \theta_{i} B^i)(1-B)^dy_t = c + (1 + \sum_{i = 1}^{q} \phi_{i} B^i)e_t \label{eq2} \\ \notag 
\] where c is a constant and \{\(e_t\)\} is a white noise process with
zero mean and variance \(\sigma^2\).

ARIMA models have commonly been used in financial forecasting and are
popular for observing stock prices and exchange rates due to its power
and statistical properties (C.-S. Lin, Chiu, and Lin 2012). They have
frequently been used as a benchmark to compare new forecasting
techniques that have emerged over time. Kamruzzaman and Sarker (2003)
applied an ARIMA model to forecast the exchange rate between the
Australian dollar and six other currencies, and used the forecast errors
as well as the accuracy of the direction of the forecasts to evaluate
the performance of three Neural Network models. In a similar study,
Khashei, Bijari, and Ardali (2012) compared the predictive capability of
their proposed hybrid model consisting of an ARIMA and a Probabilistic
Neural Network (PNN) to the traditional ARIMA model, and justified the
use of the ARIMA as a benchmark by claiming that it is the most
important linear model. Their comparison was made by investigating the
forecasts between the British pound and US dollar over varying forecast
horizons. Not only has the ARIMA model been useful as a benchmark, but
it has also yielded satisfactory results when predicting exchange rates.
When Nwankwo (2014) forecasted the rate between the Nairo and dollar,
diagnostic testing revealed that the ARIMA(1,0,0) model was the best fit
for the data based on Akaike's Information Criterion (AIC).

The ARIMA model is attractive as it is tractable and produces good
short-term forecasts when more than 100 observations are used (Tseng et
al. 2001). Although the ARIMA model has the advantage of ease of
implementation and flexibility, it fails to capture the non-linearity
and volatility present in exchange rate data. Over time, the ARIMA model
has evolved to cater for a wider variety of data and to compensate for
some of its shortcomings. The most popular versions of the ARIMA model
that has been implemented in exchange rate forecasting is the Seasonal
ARIMA (SARIMA) and Fractional ARIMA model (FARIMA).

The SARIMA model was introduced to capture the periodic behaviour of
data and extends the ARIMA class by including a term for seasonal
differencing. Etuk, Wokoma, and Moffat (2013) modelled the Naira/CFA
Franc exchange rate which exhibited monthly seasonality using an
additive SARIMA model, to demonstrate that it can be a useful fit for
exchange rate data which displays seasonality. Their results showed that
the SARIMA model adequately described the variation in the exchange rate
series. The ARFIMA model generalises the ARIMA model in that the degree
required to make the data stationary can assume any real value, and is
no longer restricted to the integer domain. The ARFIMA model has the
ability to capture the dependence between observations that are widely
spread apart in time (Cheung 1993). This makes the model more
parsimonious since it can capture long memory in data as well as short
term dynamics (Cheung 1993). Cheung (1993) fitted the ARFIMA model to
examine five exchange rates, and found that there was strong evidence of
long memory in the exchange rate time series.

Generalised autoregressive conditional heteroskedasticity (GARCH) models
were later developed in an attempt to capture the volatility in
financial markets (Anastasakis and Mort 2009). Hsieh (1989) applied a
GARCH model to investigate five exchange rates and his results showed
that although the GARCH model outperformed the random walk, some
non-linear information still remained in the residuals.

\subsection{Other techniques used to forecast exchange
rates}\label{other-techniques-used-to-forecast-exchange-rates}

Over time, financial forecasting methods have moved away from linear
models like ARIMA and GARCH, to soft computing techniques. These more
complex techniques are non-linear and can fit complex time series more
easily (Castillo and Melin 2002). Unlike the ARIMA model, soft computing
techniques do not impose structural assumptions on the model apriori
(Castillo and Melin 2002). Some of the most commonly used artificial
intelligence methods used to forecast exchange rate data are Neural
Networks and Fuzzy Logistic Systems.

Artificial Neural Networks (ANNs) have had many successful applications
in forecasting exchange rates and are more advantageous than other
non-linear forecasting methods. They are data driven, can adapt to
non-stationary environments and can approximate any continuous function
(Khashei and Bijari 2011). In a study done by Fahimifard et al. (2009),
the ANN was found as an effective way to improve the forecasts of
exchange rates. Superior results were produced when compared to the
ARIMA and GARCH model using the root mean square error (RMSE), mean
square error (MSE) and mean absolute difference as a measure of
performance. Fuzzy Logistic Systems (FLSs) were initially developed to
solve problems involving linguistic terms, and have been successfully
used in financial forecasting (Khashei, Bijari, and Ardali 2009). Fuzzy
logic tries to imitate human reasoning and the decision-making process
and allows for finer rather than discrete decisions to be provided.
Santos, Costa, and Santos Coelho (2007) investigated how well FLSs and
ANNs perform compared to the traditional ARMA and GARCH model. He
examined the forecasts of Brazilian exchange rate returns by considering
different frequencies of the series and comparing their one step-ahead
forecasts. By analysing accuracy statistics, he found that FLSs and ANNs
achieved higher returns based on the forecasts they produced. Similar
results were found by Khashei, Bijari, and Ardali (2009) when he
analysed the predictive capabilities of FLSs, ANNs, the traditional
ARIMA model, and a Fuzzy ARIMA model.

Although ANNs have been broadly applied in financial forecasting, the
process of building them is a complex task and there is no consistent
method of design compared to the traditional Box-Jenkins ARIMA model.
Unlike ARIMA models, the performance of ANNs is sensitive to many
modelling factors such as the number of input nodes included and the
size of the training sample chosen (Zhang and Hu 1998). Like ANNs there
is no systematic approach to designing FLSs and they are only
understandable when simple. Although FLSs has the advantage over ARIMA
models that they can be applied to data with few observations available,
it gives acceptable rather than accurate results and are more suitable
for problems which do not require high accuracy.

\subsection{Hybrid ARIMA Models}\label{hybrid-arima-models}

Over time, many researchers began to think of ways in which to harness
the advantages of tractable linear models such as the ARIMA model, and
of more complex non-linear models. By combining the ARIMA model with
other forecasting methods, the advantages of both forecasting methods
are leveraged while simultaneously improving their limitations. Some of
the hybrid models which have commonly been used in exchange rate
forecasting are the Fuzzy ARIMA and ANN-ARIMA model.

The ARIMA model produces very accurate forecasts over short time
horizons however it has the limitation of requiring more than 100
observations of historical data to yield accurate results (Tseng et al.
2001). In a world that is constantly changing and with the rapid
advancement of technology, access to large amounts of historical data is
difficult to obtain. On the contrary, a fuzzy regression model requires
little historical data however produces wide prediction intervals if
extreme values are present in the data. The Fuzzy ARIMA model combines
the ARIMA and fuzzy regression model to exploit the advantages of both
models while simultaneously overcoming their limitations. Tseng et al.
(2001) proposed applying a Fuzzy ARIMA model to forecast the exchange
rate of Taiwan dollars to US dollars to demonstrate the model's
appropriateness and power. The Fuzzy ARIMA not only produced forecasts
that were superior to the ARIMA and fuzzy time series models but also
provided an upper and lower bound which can be used by decision makers
to determine the best and worst possible situations.

Although ARIMA models are powerful, they require non-stationary data to
be differenced and impose prior assumptions onto the distribution of the
data (Ince and Trafalis 2006). In contrast, machine learning techniques
such as ANNs do not impose any assumptions onto the data generating
process however, being data-driven, are sensitive to the number of input
nodes used. The ANN-ARIMA model draws on the strengths of both models
and overcomes these individual difficulties. Ince and Trafalis (2006)
created an ANN-ARIMA model by using the ARIMA model to determine the
number input nodes required by an ANN for three exchange rates. When the
ANN-ARIMA model was compared to the pure ARIMA model, the hybrid model
outperformed the ARIMA based on the MSE.

Although the traditional ARIMA models produce less superior forecasts
than its hybrid forms and other complex non-linear techniques, its
forecasts are still satisfactory. They are simple models that are easy
to implement and have a consistent method of model design and selection.
ARIMA models are also more robust and efficient than complex structural
models in relation to short-run forecasting. The fact that they have
been used as the foundation for more advanced models and have commonly
been used as a benchmark for comparison, justifies it as a good starting
point to compare it to Facebook's forecasting method, Prophet, that was
recently released.

\subsection{Forecasting with Prophet}\label{forecasting-with-prophet}

The techniques that have been considered for exchange rate forecasting
thus far, require the analyst to have vocational knowledge about time
series. Prophet differs to traditional time series models in that it is
flexible and can be customised by a large number of non-experts who have
little knowledge about time series, however have domain knowledge about
the data generating process. Prophet allows for a large number of
forecasts to be produced across a variety of problems and consists of a
robust evaluation system that allows for a large number of forecasts be
evaluated and compared. This is Facebook's definition of forecasting at
scale. Prophet is similar to a Generalized Additive Model (GAM) - an
additive regression model that consists of non-linear and linear
regression functions applied to predictor variables (Taylor and Letham
2017). The decomposable model is of the form:

\[ 
    y(t) = g(t) + s(t) + h(t) + e_t
\] where the components of the model represent the growth, seasonality
and holiday respectively, and \(e_t\) is white noise.

Prophet, like the GAM, frames the forecasting problem as a curve fitting
exercise and uses backfitting to find the regression functions. This
allows for the model to be fitted quickly and missing values and large
outliers to be handled elegantly. The regression model also provides
model flexibility and allows the analyst to interactively change model
parameters (Taylor and Letham 2017).

The growth component of Prophet may be modelled as a linear or
non-linear function of time. Linear growth is modelled by a piecewise
constant function while non-linear growth is modelled similar to
population growths which use a logistic growth model (Taylor and Letham
2017). A time-varying upper limit may be specified for logistic growth,
at which point the forecasts will saturate. This carrying capacity
allows the analyst to incorporate their prior knowledge about the
maximum obtainable growth level such as the total market or population
size into the model. Prophet accounts for changes in the trajectory of
this trend by automatically detecting and selecting changepoints in the
data at which the growth rate is allowed to change. These changepoints
have a Laplace prior distribution placed on them and its scale parameter
may be used to adjust the flexibility of the trend and to choose how
aggressively the model should follow historical trend changes. Analysts
may also adjust the number of potential changepoints included or
manually specify their location. This allows non-experts with knowledge
about events that may affect the growth rate to use the parameter as a
knob to either increase or decrease the number of changepoints included
(Taylor and Letham 2017). It also allows for the analyst to add
changepoints which the automatic selection procedure may have missed or
remove changepoints when the model is overfitting historical trends
(Taylor and Letham 2017).

The decomposable form of the model allows for multiple seasonality
components with different periods to be added to the model. Seasonality
components are modelled by a Fourier series and have a Normal prior
distribution placed on its parameters (Taylor and Letham 2017). The
spread parameter can be adjusted by analysts to smooth the seasonality
and change how much of historical seasonality is projected into the
future (Taylor and Letham 2017).

The analyst may also provide a list of important events and holidays
which have impacted the time series in the past or which they know might
impact it in the future. The list could include the name, date, and
country in which they have taken place or are expected to take place
(Taylor and Letham 2017). By specifying the country of occurrence,
separate lists can be populated for global events and holidays, and
country-specific events and holidays. The union of the global and
country-specific lists can then be used for forecasting. Like
seasonality, a Normal prior distribution is placed on the parameters of
the holiday component, and the scale parameter can be adjusted by
analysts to smooth the holidays (Taylor and Letham 2017).

Prophet's Bayesian approach to forecasting allows the analyst to
incorporate their expert knowledge into the model building process and
has produced significantly improved forecasts compared to the ARIMA
model. Taylor and Letham (2017) forecasted the number of events on
Facebook using Prophet. The time series was impacted by holidays, had
strong multi-period seasonality, and a piecewise trend. The forecasts
produced by Prophet were compared to common forecasting techniques such
as exponential smoothing, ARIMA, the seasonal naïve, and the naïve
model. While exponential smoothing and the seasonal naïve model were
quite robust, the ARIMA forecasts were fragile. No model besides Prophet
accounted for the dips around holidays and the upward trend of the time
series towards later observations.
